Improvement of crash forces in structures using optimization tools.

### 2016-07-27 paper.tex
Avances

\documentclass[cmfonts]{witpress}
\usepackage{todonotes}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{cleveref}
\bibliographystyle{witpress}

@@

\author{L. E. Romera, M. Costas, J. Paz, J. D\'iaz, S. Hern\'andez.}

@@

\section{Introduction}
Security is a key issue on vehicle design nowadays, and crashworthy elements are required to absorb the energy during a crash and minimize the effects on the occupants. Crash boxes are devices equipped by vehicles that are designed to absorb the impact through deformation in order to protect the vehicle passengers. They are usually tube shaped and made of several pieces that require bonding. This bonding can be also a valid dissipation mechanism, apart from being a way of obtaining low-weight unions.

On the other hand, structural adhesives have improved their mechanical properties in recent years and now can be used with confidence in heavy duty applications. Adhesive unions present themselves as a promising solution for bonding manufacture, thanks to their properties and last years improvements on these. Interest on them is also raising up in scientific fields thanks to these enhancements and to the recent development of validated numerical models.
\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{./figures/IMG_CUTRES/comparison}
	\caption[Qualitative strength and cost comparison of different joining solutions.]{Qualitative strength and cost comparison of different joining solutions. Taken from \cite{superyacht}.}
	\label{fig:comparison}
\end{figure}

As \cref{fig:comparison} illustrates, adhesives are very competitive both in bonding strength and in economical manufacturing costs \cite{superyacht}. Although other solutions, like welding, may compete in strength, adhesives are much more inexpensive than other solutions. It should be recalled that this figure only aims to give a qualitative idea of this comparison for a particular manufacturer's point of view, and should not be taken as a fact.

In order to check adhesive's suitability for their use in crashworthy elements, this study develops numerical models of adhesively bonded crash boxes subjected to impact load  using the FEM. Information and data present in the literature are used to create and validate these models in order to ensure the accuracy of the results.

The present study aims to:

\begin{itemize}
	\item Determining an efficient adhesive modellization for its use in numerical models of adhesive joints. This modellization would also serve for studying hybrid bonding which implies the use of adhesives with other bonding solutions at the same time, such as welding or riveting.

	\item Checking the suitability of adhesives for their use in crash boxes or as a manner of improving other union types performance in that same devices.

	\item Generating a parametric model that would serve for sensitivity analysis and  optimization of the tube geometry by using the scripting capabilities of the finite element package employed.
\end{itemize}

In this introductory chapter, the most remarkable adhesive properties of interest are commented, along with experimental and numerical analysis that have been performed in the literature, with a later focus on their use in the field of crash boxes.

The present document is structured in three more chapters beyond this point, explaining the realized model, discussing the obtained results and drawing some conclusions, respectively. Two appendices are added at the end including some other aspects, such as the Python code developed for this study or the manufacturer's catalog information regarding to the selected adhesive.

\section{Finite element model}
The modelled crash box consists of two adherend plates made of steel that create a tube of square section by being bonded together with adhesive. The union is practiced on the surface of two flanges on each adherend side. Adhesive failure and adherend plastification are taken into account.

@@

\section{Geometry}

@@

\subsection{Triggering}

@@

\subsection{Impact conditions}  % Speed, etc
\label{sec:impact}

@@

\subsubsection{Rivets}

@@

The spot weld was made up of steel, using its elastic modulus as stiffness and the steel plastification stress on the normal direction as the failure criterion. Several sizes for the spot-weld points were tried up to an average solution which wasn't either too strong to obviate the adhesive's contribution, neither too weak to not prevent the critical failure.

\subsubsection{Stabilizing box}

@@

\section{Materials}

@@

\subsection{Adhesive}
Loctite Hysol 9514 \cite{manufCatalog} was the chosen adhesive for the bonding, for its shown interest by many other authors for structural bonding \cite{Sadowski2010, Scattina2011, SernaMoreno2015}, making it an \textit{a priori} good candidate for the studied application. It is a monocomponent epoxy adhesive which requires heat curing processes in the union cast.

@@

\subsubsection{Constitutive behaviour}
\label{sec:elastic}

@@

\subsubsection{Damage model}
\label{sec:damage}

@@

As long as the condition exposed in \cref{eq:quads} is not satisfied, the adhesive layer behaves elastically. Once the condition gets satisfied, damage starts and a degradation in stiffness can be appreciated following the especified damage evolution model, being thus a case of LEFM, as the traction behaviour is also linear and no plastification takes place at any point.

@@

XFEM-based LEFM require a pre-existing crack in the model or some nucleation formulation that governs its propagation \cite{Abaqus613Manual}. As the crack's initial localization and direction is not previously known, the VCCT criterion is used, which consists on correlating the displacement energy to the suffered damage. This technique can be applied in brittle crack formation, which is the case, as no plastification happens before damage or failure.

@@

\subsection{Adherends}

@@

\subsection{Interaction}

@@

\section{Computational and numerical considerations}
% Include computational time

\subsection{Hardware and software}

@@

\subsection{Model parametrization}
\label{sec:script}

@@

A flowchart explaining the model creation process of the code is depicted in \cref{fig:flow}. The whole code can be seen in \cref{apx:script}. Complementary scripts were also developed for its work within the cluster and for its optimization using the Dakota optimization software package \cite{dakota}.

@@

\subsection{On the XFEM and the explicit solver}
\label{sec:xfem}

@@

\begin{equation}
S\left(t_{i+1}\right) = F_{i}\left(S\left(t_{i}\right)\right) ;
\label{eq:explicit}
\end{equation}
meaning that a state, $S$, at a certain time step, $t_{i+1}$, is calculated using the state at the previous time step, $t_{i}$. This method updates the stiffness matrix (included with other terms into $F$) after calculating each time step in order to take into account the effects of plastification, damage, etc.

@@

\subsection{Finite element mesh}

@@

% Total recount of elements (by part?)
\begin{table}
	\centering
	\begin{tabular}{ll p{3cm} rr}

		Part & Element type & Mesh type & Elements & Nodes \\

		Adherend & S4R (S3) & Structured (with free partitions) & $3844$ ($33$) & $4010$ \\
		Adhesive & COH3D & Structured & $1200$ & $2718$ \\
		Rigid plates & R3D4 & Free & $10$ & $15$ \\
		Stabilizing box & R3D4 & Structured & $448$ & $480$ \\

		Total &&& $10622$ & $13966$ \\

	\end{tabular}
	\caption[Mesh summary.]{Mesh summary. Given values correspond to each instance of that part. In the adherend case, there is a seven element recount difference between each instance, probably due to the free mesh control in some partitions. The average value is given.}
	\label{tab:mesh}
\end{table}

\subsubsection{Mass scaling}
\label{sec:mass_scaling}

>>>>

\documentclass[cmfonts]{witpress}
\usepackage{todonotes}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{cleveref}
\bibliographystyle{witpress}
% \usepackage{natbib}

@@

\author{L. E. Romera, L. Pire, M. Costas, J. Paz, J. D\'iaz, S. Hern\'andez.}

@@

\section{Introduction}
Occupant safety is a key issue on vehicle and aircraft design nowadays, and crashworthy elements are required to absorb the energy during a crash and minimize the effects on the occupants. Aircraft and vehicles are nowadays provided with crash boxes to absorb the impact energy through deformation in order to protect the passengers. They are usually tube-shaped and made of several parts that require some kind of bonding.

On the other hand, structural adhesives have improved their mechanical properties in recent years and can be now used with confidence in heavy duty applications. Interest on them is also raising up in scientific fields thanks to these enhancements and to the recent development of validated numerical models.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{./figures/IMG_CUTRES/comparison}
	\caption[Qualitative strength and cost comparison of different joining solutions.]{Qualitative strength and cost comparison of different joining solutions. Taken from \cite{superyacht}.}
	\label{fig:comparison}
\end{figure}

As shown in \Cref{fig:comparison}, adhesives are very competitive both in bonding strength and in economical manufacturing costs \cite{superyacht}. In order to check adhesive's suitability for their use in crashworthy elements, this research develops numerical models of adhesively bonded crash boxes subjected to impact loads using finite element models. Information and data present in the literature are used to create and validate these models in order to ensure the accuracy of the results.

\section{Finite element model}
The modeled crash box consists of two adherend plates made of steel that create a tube of square section by being bonded together with an epoxy adhesive. The union is practiced on the surface of two flanges on each adherend side. Adhesive failure and adherend yielding are taken into account.

@@

\subsection{Geometry}

@@

\subsubsection{Triggering}

@@

\subsubsection{Impact conditions} % Speed, etc
\label{sec:impact}

@@

@@

The spot weld was made up of steel, using its elastic modulus as stiffness and the steel yielding stress on the normal direction as the failure criterion. Several sizes for the spot-weld points were tried up to an average solution which wasn't either too strong to obviate the adhesive's contribution, neither too weak to not prevent the critical failure.

@@

\subsection{Materials}

@@

\subsubsection{Adhesive}
Loctite Hysol 9514 \cite{manufCatalog} was the chosen adhesive for the bonding, for its shown interest by many other authors for structural bonding \cite{Sadowski2010, Scattina2011, SernaMoreno2015}, making it an \textit{a priori} good candidate for the studied application. It is a monocomponent epoxy adhesive which requires heat curing processes in the union cast.

@@

@@

@@

As long as the condition exposed in \cref{eq:quads} is not satisfied, the adhesive layer behaves elastically. Once the condition gets satisfied, damage starts and a degradation in stiffness can be appreciated following the especified damage evolution model, being thus a case of LEFM, as the traction behaviour is also linear and no yielding takes place at any point.

@@

XFEM-based LEFM require a pre-existing crack in the model or some nucleation formulation that governs its propagation \cite{Abaqus613Manual}. As the crack's initial localization and direction is not previously known, the VCCT criterion is used, which consists on correlating the displacement energy to the suffered damage. This technique can be applied in brittle crack formation, which is the case, as no yielding happens before damage or failure.

@@

\subsubsection{Adherends}

@@

\subsubsection{Interaction}

@@

\subsection{Computational and numerical considerations}
% Include computational time

\subsubsection{Hardware and software}

@@

\subsubsection{Model parametrization}
\label{sec:script}

@@

A flowchart explaining the model creation process of the code is depicted in \cref{fig:flow}. Complementary scripts were also developed for its work within the cluster and for its optimization using the Dakota optimization software package \cite{dakota}.

@@

\subsubsection{On the XFEM and the explicit solver}
\label{sec:xfem}

@@

\begin{equation}
S\left(t_{i+1}\right) = F_{i}\left(S\left(t_{i}\right)\right) ;
\label{eq:explicit}
\end{equation}
meaning that a state, $S$, at a certain time step, $t_{i+1}$, is calculated using the state at the previous time step, $t_{i}$. This method updates the stiffness matrix (included with other terms into $F$) after calculating each time step in order to take into account the effects of yielding, damage, etc.

@@

\subsubsection{Finite element mesh}

@@

% Total recount of elements (by part?)
\begin{table}
	\centering
	\begin{tabular}{ll p{3cm} rr}
		Part & Element type & Mesh type & Elements & Nodes \\
		Adherend & S4R (S3) & Structured (with free partitions) & $3844$ ($33$) & $4010$ \\
		Adhesive & COH3D & Structured & $1200$ & $2718$ \\
		Rigid plates & R3D4 & Free & $10$ & $15$ \\
		Stabilizing box & R3D4 & Structured & $448$ & $480$ \\
		Total &&& $10622$ & $13966$ \\
	\end{tabular}
	\caption[Mesh summary.]{Mesh summary. Given values correspond to each instance of that part. In the adherend case, there is a seven element recount difference between each instance, probably due to the free mesh control in some partitions. The average value is given.}
	\label{tab:mesh}
\end{table}
